---
title: "Project - 02"
author: "Binish Kurian Chandy"
date: "3/10/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RCurl)
library(tidyverse)
```

## Dataset 1 - Total health spending per person (US$)  
[Dataset 1 Location](https://www.gapminder.org/data/)  
The csv file is hosted in github for processing.    
[Dataset 1 CSV](https://raw.githubusercontent.com/binishkurian/DATA-607/master/project-02/health_spending_per_person.csv)

### Create data frame from CSV
```{r}
library(RCurl)
url <- getURL("https://raw.githubusercontent.com/binishkurian/DATA-607/master/project-02/health_spending_per_person.csv")
df <- read.csv(text = url, header = TRUE, check.names = FALSE)
```

### Tidy data
```{r}
(health_spending <- as.tibble(df) %>%
        gather(`1995`:`2010`, key = "year", value = "USD", na.rm = TRUE))
```

### Find the maximum spending per person for each country
```{r}
(max_spending_for_each_country <-health_spending %>%
        group_by(country) %>%
        summarise(max_spending = max(USD, na.rm = TRUE))) 
```

### Find the countries with maximum spending per person
```{r}
max_spending_for_each_country %>%
        mutate(r = min_rank(desc(max_spending))) %>% 
        arrange(r)
```

```
United States has most spending followed by Luxembourg, Norway, Switzerland and Monaco
```

### Let's explore the spending per person for only above countries
```{r}
health_spending %>%
        filter(country %in% c("United States", "Luxembourg", "Norway", "Switzerland", "Monaco")) %>%
        ggplot(mapping = aes(x = country, y = USD)) +
        geom_boxplot()
```

### Let's check the general trend in spending for above countries
```{r}
health_spending %>%
        filter(country %in% c("United States", "Luxembourg", "Norway", "Switzerland", "Monaco")) %>%
        ggplot(mapping = aes(x = year, y = USD)) +
        geom_line(aes(group = country), colour = "grey50") + 
        geom_point(aes(colour = country))
```
```
The general trend for United States is to have increased spending year by year.
```

## Dataset 2 - Betsy Rosalen's Movie Ratings Dataset
[Dataset 2 Location](https://docs.google.com/spreadsheets/d/1K_a87dxLQR91GwVcqkqZK9kjxcYDl1tbK9EX9WvC7bo/edit#gid=601448852)  
The csv file is hosted in github for processing.    
[Dataset 2 CSV](https://raw.githubusercontent.com/binishkurian/DATA-607/master/project-02/movie_rating.csv)

### Create data frame from CSV
```{r}
library(RCurl)
url <- getURL("https://raw.githubusercontent.com/binishkurian/DATA-607/master/project-02/movie_rating.csv")
df <- read.csv(text = url, header = TRUE, check.names = FALSE)
```

### Tidy data
```{r}
(tidy_rating <- as.tibble(df) %>%
        select(-Timestamp) %>%
        gather(`Guardians of the Galaxy 2`:`Valerian and the City of a Thousand Planets`, key = movie, value = rating))
```

### Find the movie with max average rating
```{r}
tidy_rating %>%
        group_by(movie) %>%
        summarise(av_rating = mean(rating, na.rm = TRUE)) %>%
        mutate(r = min_rank(desc(av_rating))) %>%
        arrange(r)
```
```
Wonder Woman has the max average rating of 3.12 followed by Star Wars: 
The Last Jedi with average rating of 2.92
```

### Let's do summary computation for movies
```{r}
tidy_rating %>%
        ggplot() +
        stat_summary(mapping = aes(x = movie, y = rating),
                     fun.ymin = min,
                     fun.ymax = max,
                     fun.y = mean) +
        coord_flip()
```

### Let's find the maximum rating provided by each evaluator
```{r}
tidy_rating %>%
        group_by(Name) %>%
        summarize(max_rating = max(rating, na.rm = TRUE)) %>%
        arrange(desc(max_rating))
```

### Do summary computation for movie evaluators
```{r}
tidy_rating %>%
        ggplot() +
        stat_summary(mapping = aes(x = Name, y = rating),
                     fun.ymin = min,
                     fun.ymax = max,
                     fun.y = mean) +
        coord_flip()
```

## Dataset 3 - Aged 55+ unemployment rate (%)  
[Dataset 3 Location](https://www.gapminder.org/data/)  
The csv file is hosted in github for processing.    
[Dataset 3 CSV](https://raw.githubusercontent.com/binishkurian/DATA-607/master/project-02/health_spending_per_person.csv)

### Create data frame from CSV
```{r}
library(RCurl)
url <- getURL("https://raw.githubusercontent.com/binishkurian/DATA-607/master/project-02/unemployment_above_55.csv")
df <- read.csv(text = url, header = TRUE, check.names = FALSE)
```

### Tidy data
```{r}
tidy_unemploy <- as.tibble(df) %>%
        gather(`1981`:`2005`, key = year, value = prec, na.rm = TRUE)
```

### Find maximum unemployment percentage for each country
```{r}
tidy_unemploy %>%
        group_by(country) %>%
        summarise(max_perc = max(prec, na.rm = TRUE))
```

### Find the country with max unemployment percentage
```{r}
tidy_unemploy %>%
        group_by(country) %>%
        summarise(max_perc = max(prec, na.rm = TRUE)) %>%
        arrange(desc(max_perc))
```
```
Latvia has the maximum unemployment percentage followed by Finland.
```

### Scatter plot for unemployment percentage
```{r}
tidy_unemploy %>%
        ggplot(aes(x = year, y = prec, group = country)) +
        geom_jitter() +
        coord_flip()
```