---
title: "Object Oriented Programming in R"
author: "Binish Kurian Chandy"
date: "5/1/2018"
output: html_document
---

```{r setup, include=FALSE}
require(pryr)
knitr::opts_chunk$set(echo = TRUE)
```

### What is Object Oriented Programming?
Object-oriented programming is a programming paradigm based on the concept of "objects", 
which may contain data, in the form of fields, often known as attributes; and 
code, in the form of procedures, often known as methods.
 
A **class** defines the behavior of **objects** by describing their attributes and their 
relationship to other classes. The class is also used when selecting **methods**, 
functions that behave differently depending on the class of their input. 
Classes are usually organised in a hierarchy: if a method does not exist for a 
child, then the parent’s method is used instead; the child **inherits** behavior 
from the parent. 

### R’s three OO systems
* **S3** implements a style of OO programming called generic-function OO. This is 
different from most programming languages, like Java, C++, and C#, which implement 
message-passing OO. S3 is a very casual system. It has no formal definition of classes.

* **S4** works similarly to S3, but is more formal. **S4 has formal class definitions**, 
which describe the representation and inheritance for each class, and has special helper 
functions for defining generics and methods. 

* **Reference classes**, called RC for short, are quite different from S3 and S4.
  + Introduced in version `2.12`
  + RC methods belong to objects, not functions
  + RC objects are mutable

### Reference classes (RC)
RC objects behave more like objects do in most other programming languages, 
e.g., Python, Ruby, Java, and C#. RC classes are best used for describing stateful 
objects, objects that change over time.

### Defining classes and creating objects
```{r}
# setRefClass() generates new objects
Account <- setRefClass("Account")
Account$new()

# how to create fields ?
Account <- setRefClass("Account",
  fields = list(balance = "numeric"))

# access field using $
a <- Account$new(balance = 100)
a$balance

# mutate field 
a$balance <- 200
a$balance

# RC objects are mutable
b <- a
b$balance
a$balance <- 0
b$balance

# For this reason, RC objects come with a copy() method that allow you to make a copy of the object
c <- a$copy()
c$balance
a$balance <- 100
c$balance

# Add behaviour to objects through methods
# contains is the name of the parent RC class to inherit behaviour from.
# The deep assignment arrow, <<-, never creates a variable in the current environment,
# but instead  modifies an existing variable found by walking up the parent environments.
Account <- setRefClass("Account",
  fields = list(balance = "numeric"),
  methods = list(
    withdraw = function(x) {
      balance <<- balance - x
    },
    deposit = function(x) {
      balance <<- balance + x
    }
  )
)

a <- Account$new(balance = 100)
a$deposit(100)
a$balance
```

### Inheritence
```{r}
NoOverdraft <- setRefClass("NoOverdraft",
  contains = "Account",
  methods = list(
    withdraw = function(x) {
      if (balance < x) stop("Not enough money")
      balance <<- balance - x
    }
  )
)
accountJohn <- NoOverdraft$new(balance = 100)
accountJohn$deposit(50)
accountJohn$balance
```
```{r, eval=FALSE}
accountJohn$withdraw(200)
```
All reference classes eventually inherit from **envRefClass**.

### Recognizing objects and methods
```{r}
is(accountJohn, "refClass")
pryr::otype(accountJohn)
```

### Method dispatch
Method dispatch is very simple in RC because methods are associated with classes, 
not functions. When you call `x$f()`, R will look for a method f in the class of x, 
then in its parent, then its parent’s parent, and so on. From within a method, 
you can call the parent method directly with `callSuper(...)`.









