---
title: "NoSQL Migration"
author: "Binish Kurian Chandy"
date: "4/24/2018"
output: html_document
---

```{r setup, include=FALSE}
library(RPostgreSQL)
library(mongolite)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```
## Read movie data from PostgresSQL DB
```{r}
driver <- dbDriver("PostgreSQL")
con <- dbConnect(driver, host="localhost", port=5432, dbname="movies", user="postgres")
query <- "select m.mid, m.title, m.director, re.name, ra.stars from movie m inner join rating ra on m.mid = ra.mid inner join reviewer re on
ra.rid = re.rid"
rs <- dbSendQuery(con, query)
movies <- dbFetch(rs, n=-1)
head(movies)
```

## How to install MongoDB on mac OS?
```
Using home brew
brew install mongodb
mongod # to start mongodb
```

## Write data to MongoDB
```{r}
movies_collection = mongo(collection = "movies", db = "movies_db")
movies_collection$insert(movies)
```

## Verify count of rows from both databases
```{r}
nrow(movies)
movies_collection$count()
```

## Get the first record from mongo db
```{r}
unlist(movies_collection$iterate()$one())
```

## Relational vs. NoSQL database
`Advantages`

* `Non-Relational means table-less:` NoSQL databases are non-relational, hence, 
  very different from SQL  databases. This means they are easier to manage and 
  they provide a higher level of flexibility with newer data models.
 
* `Mostly Open Source and Low-Cost:` The open source nature of NoSQL databases 
   makes them an appealing solution for smaller organizations with limited budgets. 
   The top NoSQL databases on the market today (MongoDB, MarkLogic, Couchbase, CloudDB, 
   and Amazon’s Dynamo DB) allow for rapid processing of real-time Big Data 
   applications in ways that are affordable.
 
* `Easier scalability through Sharding:` NoSQL database experts often use elastic 
   scalability as a major selling point of NoSQL. NoSQL databases are designed to 
   function on full throttle even with low-cost hardware.
 
* `No need to develop a detailed database model:` The non-relational nature of 
   a NoSQL database allows database architects to quickly create a database without 
   needing to develop a detailed (fine-grained) database model. This saves a 
   lot of development time.
 
`Disadvantages`

* `Community not as well defined:` While its continuing to grow rapidly, the 
  NoSQL community is relatively new and lacks the maturity of the MySQL user base. 
  Obviously, NoSQL is growing rapidly, but for now MySQL is hard to beat for its 
  vast network of highly experienced end users.
 
* `Lack of reporting tools:` A major problem with NoSQL databases is the lack of 
  reporting tools for analysis and performance testing. However, with relational dbs, you 
  can find a wide array of reporting tools to help you prove your application’s validity.
 
* `Lack of standardization:` In order for NoSQL to grow, it needs a standard query 
  language like SQL.