---
title: "Assignment – SQL and R"
author: "Binish Kurian Chandy"
date: "2/8/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### How to build the postgress database?
All files used to build database can be found at :
[link](https://github.com/binishkurian/DATA-607/tree/master/week-02)

Use backup file or execute the sql commands from pgAdmin to load the database.

**Load required drivers to connect to database server.**
```{r message=FALSE}
#install.packages("RPostgreSQL")
library(RPostgreSQL)
```

**Open connection to database**
```{r}
driver <- dbDriver("PostgreSQL")
con <- dbConnect(driver, host="localhost", port=5432, dbname="movies", user="postgres")
```

**Fetch movie ratings from database.**    
The movie ratings are fetched by join query across three tables namely movie, rating and reviewer.  
  **Movie** table contains movie name, movie id and director name.  
  **Reviewer** contains reviewer id and name.  
  **Rating** contains movie id, reviewer id and stars.  
```{r message=FALSE}
query <- "select m.mid, m.title, m.director, re.name, ra.stars from movie m inner join rating ra on m.mid = ra.mid inner join reviewer re on
ra.rid = re.rid"

rs <- dbSendQuery(con, query)
movies <- dbFetch(rs, n=-1)
head(movies)
dbDisconnect(con) 
dbUnloadDriver(driver)
```

**How many movies received 5 star rating?**
```{r}
unique(movies[which(movies$stars == 5), ]$title)
```